---
title: "Objective2_Logistic_Regression"
author: "Christina Kuang"
date: "4/19/2022"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ROCR)
#turning off scientific notation
options(scipen = 99999999) 

#reading in data
Data <- read.csv("Nursing.csv")
Data <- subset(Data, select = -c(X)) #remove column X

Data$Rural <- factor(Data$Rural)
levels(Data$Rural) <- c("Non-Rural", "Rural")
#glimpse(Data)
contrasts(Data$Rural)
```

```{r}
set.seed(10) ##for reproducibility to get the same split
sample<-sample.int(nrow(Data), floor(.80*nrow(Data)), replace = F)
train<-Data[sample, ] ##training data frame
test<-Data[-sample, ] ##test data frame
```

```{r}
ggplot(Data, aes(x=NurseSalaries, color=Rural))+
  geom_density()+
  labs(title="Nurse Salaries by Rural vs. Non-rural")
```
```{r}
ggplot(Data, aes(x=Beds, color=Rural))+
  geom_density()+
  labs(title="Beds by Rural vs. Non-rural")
```

```{r}
ggplot(Data, aes(x=InPatientDays, color=Rural))+
  geom_density()+
  labs(title="In Patient Days by Rural vs. Non-rural")
```
```{r}
ggplot(Data, aes(x=AllPatientDays, color=Rural))+
  geom_density()+
  labs(title="All Patient Days by Rural vs. Non-rural")
```

```{r}
ggplot(Data, aes(x=PatientRevenue, color=Rural))+
  geom_density()+
  labs(title="Patient Revenue by Rural vs. Non-rural")
```
```{r}
ggplot(Data, aes(x=FacilitiesExpend, color=Rural))+
  geom_density()+
  labs(title="Facilities Expenditure by Rural vs. Non-rural")
```
```{r}
result <- glm(Rural~NurseSalaries+FacilitiesExpend+Beds+PatientRevenue+AllPatientDays+InPatientDays, family = "binomial", data=train)
summary(result)
```

H0: $\beta_1 = \beta_2 = \beta_3 = \beta_4 = \beta_5 = \beta_6 = 0$
Ha: at least one of the coefficients (betas) in H0 is not zero. 

Since the p-value is 0.018 and it is less than 0.05, we reject the null hypothesis. The 6 predictor model is better at prediction than the intercept-only model. 
```{r}
#delta G^2 test to see if coefficients for all predictors are 0. Null deviance in output minus residual deviance in output. 
TS <- result$null.deviance-result$deviance
TS

1-pchisq(TS,5)
```

$\beta_1 = NurseSalaries$
$\beta_2 = Beds, InPatientDays, AllPatientDays, PatientRevenue, FacilitiesExpend$

H0: predictors in $\beta_2 = 0$ 
Ha: at least one of the coefficients in beta2 is nonzero

Faile to reject the null. can remove the 5 predictors. 
```{r}
reduced <- glm(Rural~NurseSalaries, family = "binomial", data=train)
summary(reduced)

TS_2 <- reduced$deviance-result$deviance
TS_2

1-pchisq(TS_2,5)
```

```{r}
##predicted survival rate for test data based on training data
preds<-predict(reduced,newdata=test, type="response")

##produce the numbers associated with classification table
rates<-prediction(preds, test$Rural)

##store the true positive and false postive rates
roc_result<-performance(rates,measure="tpr", x.measure="fpr")

##plot ROC curve and overlay the diagonal line for random guessing
plot(roc_result, main="ROC Curve for predicting rural")
lines(x = c(0,1), y = c(0,1), col="red")
```

```{r}
##compute the AUC
auc<-performance(rates, measure = "auc")
auc@y.values
hist(preds)
```

```{r}
##confusion matrix. Actual values in the rows, predicted classification in cols
confusion <- table(test$Rural, preds>0.4)
confusion
```

```{r}
tp <- confusion[2,2]
tn <- confusion[1,1]
fp <- confusion[1,2]
fn <- confusion[2,1]
error_rate <- (fp+fn)/(fp+fn+tn+tp)
accuracy_rate <- (tp+tn)/(fp+fn+tn+tp)

```