---
title: 'STAT 6021: Project 2'
author: "Group 2"
date: '2022-05-06'
output: pdf_document
---


## Group 2 Members:
Gregory Madden, Christina Kuang, Chi Do, Trey Hamilton


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, tidy = TRUE, cache =TRUE)
library(multcomp)
library(lawstat)
library(leaps)
library(tidyverse)
library(faraway)
library(Stat2Data)
#turning off scientific notation
options(scipen = 99999999) 
```

***Description of Dataset (provided by Stat2Data Package)***

Our dataset contains characteristics of nursing homes in New Mexico, with 52 observations and 7 variables. Each row represents an individual nursing home.

***Variables/Columns: ***

Beds	Number of beds in the nursing home

InPatientDays	Annual medical in-patient days (in hundreds)

AllPatientDays	Annual total patient days (in hundreds)

PatientRevenue	Annual patient care revenue (in hundreds of dollars)

NurseSalaries	Annual nursing salaries (in hundreds of dollars)

FacilitiesExpend	Annual facilities expenditure (in hundreds of dollars)

Rural	1=rural or 0=non-rural

```{r}
#reading in data
data(Nursing)
Data<-Nursing
Data$Rural <- factor(Data$Rural)
levels(Data$Rural) <- c("Non-Rural", "Rural")
#checking categorical variable classifiers
contrasts(Data$Rural)
```

Glimpse of data overview: 
```{r}
glimpse(Data)
```

## ***Question 1.*** What characteristics of nursing homes in New Mexico dictate annual nurse salaries at those institutions? 

***Practical implications of a linear model for predicting cumulative annual nurse salaries for a given nursing home could be used by policymakers to rationally distribute subsidy funds to institutions that are expected to contribute the lowest salaries to a particular area.***

***Objective 1: Fit a multiple linear regression model with cumulative annual nurse salaries for individual nursing homes using the available financial characteristics for each institution. The goal is to develop a model using these available data to reliably predict institutions with low annual nursing salaries among the larger group of nursing homes across the state. ***


## Exploratory Data Analysis: 

Based on the scatter plots and correlation table, it appears that Nurse Salaries has a moderate correlation with Beds, All Patient Days, and Patient Revenue. There also appears to be a strong linear relationship between Beds and AllPatientDays, Beds and Patient Revenue, In Patient Days and All Patient Days, In Patient Days and Patient Revenue, and All Patient Days and Patient Revenue. 

```{r}
cor(Data[1:6])
```

```{r}
pairs(~ NurseSalaries + PatientRevenue + FacilitiesExpend + Beds + InPatientDays + AllPatientDays, data = Data, lower.panel = NULL)
```

Nursing home salaries by census in Annual total patient days (in hundreds) and Nursing home size (by number of beds). There appears to be a moderately strong correlation between annual total patient days and average nurse salary, with at least one apparent outlier in terms of high patient days and low salary, observation #26. 

```{r}
Data %>%
  ggplot(aes(x=AllPatientDays, y = NurseSalaries, size = PatientRevenue)) +
  geom_point(alpha = 0.4) +
  labs(x="Annual total patient days (hundreds)", y="Average Nurse Salary (hundreds $)", title = "Nursing Home Nurse Salaries") +
  guides(size = guide_legend(title = "Annual patient \ncare revenue\n (hundreds $)"))
```

Boxplot demonstrating differences in Institutional nurse's salaries in New Mexico for Rural Areas compared with Non-Rural: 
Based on the box plot, it appears there is a greater variability for non-rural nurse salary. The nurses in non-rural regions also have a higher median salary. 

```{r}
Data$Rural <- factor(Data$Rural)
levels(Data$Rural) <- c("Non-Rural", "Rural")

Data %>%
ggplot(aes(x=Rural, y=NurseSalaries))+
  geom_boxplot()+
  labs(x="", y="Nurse Salary", title = "Nursing Home Nurse Salaries (hundreds $)") 
```
Scatter plot of Patient Revenue versus Nurse Salaries: 
The slopes are not parallel, which indicates there is an interaction effect between Patient Revenue and Nurse Salaries. Again seen is outlying observation #26.

```{r}
ggplot(Data,aes(x=PatientRevenue,y=NurseSalaries,color=Rural))+
  geom_point()+
  geom_smooth(method ="lm", se=FALSE)+
  labs(x="PatientRevenue",y="NurseSalaries",title="PatientRevenue versus NurseSalaries")
```

## Carrying out initial automated search procedures: 

Using backward selection to find the best model according to AIC. Start with the first-order model with all the predictors.

***The model selected is:  NurseSalaries ~ PatientRevenue + FacilitiesExpend + 
    Rural***

```{r, echo=FALSE}
#intercept only
regnull <- lm(NurseSalaries~1,data=Data)
#full
regfull <- lm(NurseSalaries~.,data=Data)
#backward elimination 
step(regfull, scope=list(lower=regnull,upper=regfull),direction="backward")
```

Reduced model summary

```{r}
reduced <- lm(NurseSalaries~PatientRevenue + FacilitiesExpend + 
    Rural,data=Data)
summary(reduced)
```

## Identifying outlying observations. 

***Observation #26 is outlying.***

```{r}
## externally studentized residuals, t_i
ext.student.res<-rstudent(reduced) 
## identify outliers with t_i## 
## critical value using Bonferroni procedure
n<-dim(Data)[1]
#p is the number of predictors + 1 for the intercept
p<-4
crit<-qt(1-0.05/(2*n), n-1-p)
## identify
ext.student.res[abs(ext.student.res)>crit]
```

##  Identifying observations that have high leverage.

Calculating the leverage values $h_{ii}$ below and identifying ones that are >2*p/n.

***Observations 26 and 31 have high leverage. ***

```{r}
lev <-lm.influence(reduced)$hat
lev[lev>2*p/n]
```

***Observation # 26 is both outlying in the predictors and appears to have high leverage.*** Unfortunately, this dataset nor the primary reference [Smith et al. "A Comparison of Financial Performance, Organizational Characteristics, and Management Strategy Among Rural and Urban Nursing Facilities," Journal of Rural Health, 1992, pp 27-40.] provide identifying information for individual Nursing Home facilities, so we cannot make specific conclusions about the facility for observation #26. What we can say about observation 26 is that it is a relatively large facility with 221 beds, especially among other rural facilities. For example, the number of beds for this facility is 79% higher than the next largest rural nursing home (123 beds). Also, despite very high patient revenue and Patient census, the nurses salary is the lowest of the entire dataset. Therefore, we suspect this facility may not be comparable with other institutions, perhaps owing to it's unique combination of large facility and rural classification. Since our primary objective is to identify low nursing salaries particularly in rural areas, we proposed re-running the regression while excluding observation #26 on the basis that is is an extraordinarily large (>200 beds) rural facility. Future predictions for such institutions will not be made using this model unless rural facilities are under 200 beds. Separate policy considerations will then be made for rare large/rural institutions.

```{r}
Data <- Data[-26,]

```

Interestingly, after re-running the EDA plots, we noticed that the slope and the intercept of the relationship between PatientRevenue and NurseSalaries no longer appear to change depending on rural vs. non-rural status. This interaction effect appears gone after excluding observation #26. 

```{r}
Data %>%
ggplot(aes(x=PatientRevenue,y=NurseSalaries,color=Rural))+
  geom_point()+
  geom_smooth(method ="lm", se=FALSE)+
  labs(x="PatientRevenue",y="NurseSalaries",title="PatientRevenue versus NurseSalaries")
```

Next, we repeated the automated search procedures with backward selection again after outlier observation #26 was removed.

***The model selected is:  NurseSalaries ~ InPatientDays + AllPatientDays + FacilitiesExpend ***

```{r}
#intercept only
regnull <- lm(NurseSalaries~1,data=Data)
#full
regfull <- lm(NurseSalaries~.,data=Data)
#backward elimination 
step(regfull, scope=list(lower=regnull,upper=regfull),direction="backward")
```

### Final Reduced model summary

```{r}
reduced <- lm(NurseSalaries~InPatientDays + AllPatientDays + FacilitiesExpend,data=Data)
summary(reduced)
```


***All VIFs below 4 so multicollinearity does not appear to be an issue. ***

```{r}
#requires package faraway
vif(reduced)
```

## Checking linear regression assumptions: 

***Assumptions 1/2 appear to be generally met: Mean of Errors = 0, constant variance. ***

```{r}

#storing fitted y residuals
yhat <- reduced$fitted.values
res <- reduced$residuals
#add to data frame

Data <-data.frame(Data,yhat,res)

#residual plot
Data %>%
  ggplot(aes(x=yhat,y=res)) +
  geom_point() +
  geom_hline(yintercept=0,color="red")+
  labs(x="fitted y", y = "residuals", title="residual plot")
```

***ACF Plot: no autocorrelation seen.***

```{r}
acf(res,main="ACF Plot of Residuals")

```

***Errors are fairly normally distributed.***

```{r}
qqnorm(res)
qqline(res,col="red")
```

Using boxcox method, we see that 1 lies just within the 95% CI for lambda so we did not transform the y variable. 

```{r}
boxcox(reduced)
```

Since we will be using this model to predict nursing salaries for new data, we are interested in the $R^2_{prediction}$.

Based on this value, the final model might be able to explain 66.15% of the variability in the new observations (as long as they are not rural facilities >200 beds). The R2 is 0.7176. Both values are fairly close to each other, so overfitting is not a major concern.

```{r}
#creating function to calculate PRESS statistic
PRESS <- function(model) {
    i <- residuals(model)/(1 - lm.influence(model)$hat)
    sum(i^2)
}

#PRESS(reduced)
## Find SST
anova_result<-anova(reduced) 
SST<-sum(anova_result$"Sum Sq") 
## R2 pred 
Rsq_pred <-1-PRESS(reduced)/SST 
Rsq_pred
```

## So, our final multiple linear regression equation is: 

$NursingSalary=366.08897-6.77821*InpatientDays+15.73249*AllPatientDays+0.15552*FacilitiesExpenditure$

***Where bed size for rural institutions is <200, NursingSalary = Estimated Annual nursing salaries (in hundreds of dollars), InpatientDays represents annual medical in-patient days (in hundreds), AllPatientDays represents annual total in-patient days (in hundreds), and FacilitiesExpenditure = hundreds of $.***

## \underline{Objective 1 Conclusions:} 

1) Patient census parameters (both total inpatient days as well as medical inpatient days) and total facilities expenditures appear to be the most important factors in determining nursing salaries among the predictors we looked at (others included PatientRevenue, Rural vs. Non-Rural). 

2) Based on the $R^2_{prediction}$, our final model summarized above might be able to explain 66% of the variability in nursing salaries of future nursing homes, as long as they are not rural facilities >200 beds. 


---

```{r  include=FALSE}
library(ROCR)
library(gridExtra)

#reading in data again
Data <- read.csv("Nursing.csv")
Data <- subset(Data, select = -c(X)) #remove column X

Data$Rural <- factor(Data$Rural)
levels(Data$Rural) <- c("Non-Rural", "Rural")
#glimpse(Data)
contrasts(Data$Rural)
```

**Question 2) What characteristics of nursing homes in New Mexico help predict if a nursing home is rural or non-rural?**

***Rural patients suffer from a lack of locally available nursing home beds. Understanding the relationships of these characteristics and how they define  rural vs. non-rural nursing homes is helpful to know how to make rural nursing homes financially viable. ***


**Objective 2: Use characteristics of nursing homes variables to develop a logistic regression model that helps predict whether a nursing home is rural or non-rural**


```{r}
set.seed(10) ##for reproducibility to get the same split
sample<-sample.int(nrow(Data), floor(.8*nrow(Data)), replace = F)
train<-Data[sample, ] ##training data frame
test<-Data[-sample, ] ##test data frame
```

## Data Exploration
The density plots of nurse salaries for rural are right skewed, which means a higher proportion of rural nursing homes have lower annual nurse salaries. In contrast, the density plots of nurse salaries for non-rural is left skewed, which means a higher proportion of non-rural nursing homes have higher annual nurse salaries. 
 
Similarly, the density plots of patient revenue for rural is right skewed, which means a higher proportion of rural nursing homes have lower annual patient revenues. In contrast, the density plots of nurse salaries for non-rural is left skewed, which means a higher proportion of non-rural nursing homes have higher annual patient revenues. 
 
These two variables, nurse salaries and patient revenue, may be good predictors, because the density plots for rural and non-rural are not very similar.
 
In comparison, the density plots of beds, patient days, all patient days, and facilities expenditure are similar for rural and non-rural nursing home facilities. As a result, these variables are less likely to be good predictors for whether a nursing home is rural or non-rural. 

```{r, echo=FALSE}
g1<-ggplot(Data, aes(x=NurseSalaries, color=Rural))+
  geom_density()+
  labs(title="Nurse Salaries by Rural vs. Non-rural") +
  theme(plot.title=element_text(size=9))

g2<-ggplot(Data, aes(x=Beds, color=Rural))+
  geom_density()+
  labs(title="Beds by Rural vs. Non-rural") +
  theme(plot.title=element_text(size=9))

g3<-ggplot(Data, aes(x=InPatientDays, color=Rural))+
  geom_density()+
  labs(title="In Patient Days by Rural vs. Non-rural") +
  theme(plot.title=element_text(size=9))

g4<-ggplot(Data, aes(x=AllPatientDays, color=Rural))+
  geom_density()+
  labs(title="All Patient Days by Rural vs. Non-rural") +
  theme(plot.title=element_text(size=9))

g5<-ggplot(Data, aes(x=PatientRevenue, color=Rural))+
  geom_density()+
  labs(title="Patient Revenue by Rural vs. Non-rural") +
  theme(plot.title=element_text(size=9))

g6<-ggplot(Data, aes(x=FacilitiesExpend, color=Rural))+
  geom_density()+
  labs(title="Facilities Expenditure by Rural vs. Non-rural") +
  theme(plot.title=element_text(size=9))

grid.arrange(g1,g2,g3,g4,g5,g6, ncol = 2, nrow = 3)
```

## Full Logistic Regression Model with all predictors
```{r}
result <- glm(Rural~NurseSalaries+FacilitiesExpend+Beds+PatientRevenue+AllPatientDays+InPatientDays, family = "binomial", data=train)
summary(result)
```

## Wald Test for $\beta_1$

$H_0:\beta_1 = 0$  
$H_a:\beta_1 \ne 0$  

With a p-value of 0.04, we can reject the null hypothesis. There appears to be a significant relationship between $\beta_1$ and the response variable. 

## 95% onfidence interval for $\beta_1$

The 95% confidence interval for $\beta_1$ is (-0.001413439498, -0.000005500502). In other words, we are 95% confident the odds of a nursing home being rural is between (exp -0.001413439498, exp -0.000005500502) = (0.9985876, 0.9999945) times the odds of a nursing home being non-rural, for given value of other predictors. Since 0 does not lie within the confidence interval, there appears to be a significant effect of nurse salaries on whether a nursing home facility is rural or non-rural, for given values of other predictors. This is consistent with our Wald Test results $\beta_1$.


```{r}
n <- 41
p <- 7
se_beta1 <- 0.00034640
beta1 <- -0.00070947
critical <-qt((1-(0.05/2)),(n-p))
multiplier <- critical*se_beta1
lower_bound <- beta1 - multiplier
upper_bound <- beta1 + multiplier
print(c(lower_bound,upper_bound))
```

## Delta G-Squared Test to see if we can drop predictors (Beds, InPatientDays, AllPatientDays, PatientRevenue, and FacilitiesExpend) that have an insignificant Wald test.

$\beta_1 = NurseSalaries$  
$\beta_2 = Beds, InPatientDays, AllPatientDays, PatientRevenue, FacilitiesExpend$  

$H_0:$ predictors in $\beta_2 = 0$   

$H_a:$ at least one of the coefficients in $\beta_2$ is nonzero  

With a p-value of 0.35, we fail to reject the null. Since none of the subset predictors appear to be significant, we can drop them from our model. 
```{r}
reduced <- glm(Rural~NurseSalaries, family = "binomial", data=train)
summary(reduced)

TS <- reduced$deviance-result$deviance
TS

1-pchisq(TS,5)
```

## Delta G-squared Test to see if our model is better than the intercept-only model
$H_0$: $\beta_1 = 0$  
$H_a$: $\beta_1 \ne 0$ 

The test statistic is 8.53 with a p-value of 0.014. Since the p-value is less than 0.05, we can reject the null hypothesis. The model is better at prediction than the intercept-only model. 
```{r}
#delta G^2 test to see if coefficients for all predictors are 0. Null deviance in output minus residual deviance in output. 
TS_2 <- reduced$null.deviance-reduced$deviance
TS_2

1-pchisq(TS_2,2)
```

## Final Model
$log(\pi/(1-\pi))$= 3.0935426 -0.0006423(NurseSalaries)

## Validate model using testing data

Since the ROC curve is above the diagonal line and AUC value is greater than 0.5, it tells us that the logistic regression model performs better than random guessing. 
```{r}
##predicted survival rate for test data based on training data
preds<-predict(reduced,newdata=test, type="response")

##produce the numbers associated with classification table
rates<-prediction(preds, test$Rural)

##store the true positive and false postive rates
roc_result<-performance(rates,measure="tpr", x.measure="fpr")

##plot ROC curve and overlay the diagonal line for random guessing
plot(roc_result, main="ROC Curve for predicting rural")
lines(x = c(0,1), y = c(0,1), col="red")
```

```{r}
##compute the AUC
auc<-performance(rates, measure = "auc")
auc@y.values
```

## Confusion Matrix

With a threshold of 0.4, the error rate is 0.09 and the accuracy rate is 0.91. In addition, the false positive rate is 0.33 and the false negative rate is 0. We decided to set the threshold to 0.4, because we are more concerned with the false negative rate. We want a lower false negative rate, because we don’t want to classify a nursing home as non-rural when it is indeed rural. If we incorrectly classify a rural nursing home facility as non-rural, it may affect the facilities' overall public funding. It is also important to note that since the dataset is small with an unblanaced number of rural and non-rural nursing homes.

```{r}
##confusion matrix. Actual values in the rows, predicted classification in cols
confusion <- table(test$Rural, preds>0.4)
confusion
```

```{r}
tp <- confusion[2,2]
tn <- confusion[1,1]
fp <- confusion[1,2]
fn <- confusion[2,1]
fpr <- fp/(tn+fp)
fpr
fnr <- fn/(fn+tp)
fnr
error_rate <- (fp+fn)/(fp+fn+tn+tp)
error_rate
accuracy_rate <- (tp+tn)/(fp+fn+tn+tp)
accuracy_rate
```

## \underline{Objective 2 Conclusions:}

**Based on our results, we conclude that nursing  salaries appear to be the most important factor in determining if a nursing home is rural or non-rural. This information can be used by policymakers to close the financial gap between rural and non-rural nursing homes.**
**For each additional $100 in annual nurse salary, the log odds of being a rural nursing home facility decreases by 0.00064.**


